title: "Homework 3"
editor: Peter Nuñez
freeze: auto
published-title: "Due date"
date: 2024-06-02
date-modified: 
---


## Set up

GitHub steps:  

1. Fork the [workshop 8 repo on GitHub](https://github.com/an-bui/ENVS193DS_workshop-08).  
2. Rename the repository to `lastname-firstname_homework-03` (use your own last name and first name!).  

RStudio/code steps:  

3. Clone your repository to your computer.  
4. Create a new Quarto document for your homework submission in the `code` folder.  
5. At the top of your document, include a link to _your forked repository_.  
6. Do all of your set up (reading in packages/data) _at the top of your document._  

Throughout the course of doing your homework, make at least 10 commits and pushes to the remote.  

## Problems

### Problem 1. Multiple linear regression: model selection and construction (52 points)

Use the information from the `homework-starter-doc.qmd` to do this problem.


#### a. Make a table _or_ list of all the models from class and the last one you constructed on your own. Write a caption for your table. (8 points)  

**Caption: Table 1:Comparing 5 models that describe total biomass. The table is comparing how well each model and it's variables describes total biomass. Model 4 best describes total biomass because it has the lowest AIC, and conforms to assumptions of linear models. ** 
table captions typically go above the table. Number the table and provide a title. Describe what is in the table (columns and rows).  
```{r model selection}
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)

par(mfrow = c(2, 2))
plot(model1)

```


```{r model-selection}
model.sel(model0,
          model1, 
          model2, 
          model3,
          model4)     #comparing the 4 models to find the one that best predicts total biomass using an Akaike Information Criterion table
```

**Table:** In your table, each row should be a model with the model number (1, 2, 3, etc.) and the predictors for each model.  

:::{.callout-note}
#### Double check your work!
There should be 5 models total (null model + 4 models with predictors).
:::

#### b. Write a 5-6 sentence "statistical methods" section. (8 points)

**To examine how specific leaf area, water treatment, and species influences total biomass, I first plotted how each variable indivudlally affected total biomass. From this initial analysis, I determined that SLA, water treatment, and species all have some effect on total biomass. I then created 5 models; one null model(no variables considered), one saturated model(all three varibale considered), and three model with two of the three variables considered. To determine the model that best described total biomass, I compared all 5 models using an AIC (Akaike Information Criterion) table. From this table, I chose the final model to be Model 4 (the model that analyzed total biomass as a function of water treatment and species) because it had the lowest AIC (-156). To evaluate the linear model assumptions, I looked at the diagnostic plots and concluded that this model adequetly follows the linear model assumptions.   **

#### c. Make a visualization of the model predictions with underlying data for your "best" model. (20 points)  

Show and annotate all your code. For full credit:  

- make the underlying data more transparent than the model predictions  
- display species names in full (not as species codes like ENCCAL or ESCCAL)  
- display water treatment types in full (not as WW or DS)  
- represent well-watered and drought stressed treatments with different colors  
- use colors that are _not_ the default `ggplot()` colors  
- facet your plot by species  
- remove the legend  
- finalize the plot  

```{r model predictions}
model_predictions <- ggpredict(model4,  #view the predictions as a data frame
                         terms = c("water_treatment", 
                                   "species_name"))   #specifing what terms we want in the plot
```

```{r model predictions plot}
plot(model_predictions, # model predictions
     limit_range = TRUE, # limit the range of predictions to the range of predictor values
     show_data = TRUE) + # show the underlying data
   facet_wrap(~group) + #grouped the plot by species
  theme_classic() + # classic theme
  theme(legend.position = "none") + #removing the legend
  labs(title = "Model 4 visualization") + # plot title
  theme(panel.grid = element_blank()) +   #removing the gridlines
  scale_color_manual(values = c("red", "blue", "darkgreen", "purple", "pink", "orange", "gray")) #using colors that aren't default ggplot colors
```
**#COME BACK TO THIS**
  custom_colors <- c("A", "B", "C", "D", "E", "F", "G")

:::{.callout-note}
Make sure that the only output is the visualization!
:::

#### d. Write a caption for your visualization. (6 points)  



#### e. Write a 3-4 sentence results section. (10 points)

Your answer should be in paragraph form and address the following points:  

- what predictors "best" described total mass (include model statistics here)?  
- on average, what differences did you find between water treatments?  
- on average, what differences did you find between species?  

### Problem 2. Affective visualization (24 points)

In this problem, you will create an **affective visualization** using your personal data in preparation for workshop during week 10.  

In lecture, we talked about the three vertices of data visualization: 1) exploratory, 2) affective, and 3) communicative. We've done a lot of exploratory and communicative visualization, but have yet to think about affective visualization.  

**Before starting, update your spreadsheet of observations.**  

#### a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences). (2 points)

**My affective visualization for my personal data could function as a psuedo tile plot. I would lay out all of my clothes, organized by their color. Meaning that they would be organized in rows, and grouped by color. I would then take a video of the clothes layed out on the floor, for a visualization of the diversity of color in my closet.  **

#### b. Create a sketch (on paper) of your idea. (2 points)  

Include a photo of this sketch in your submission.  

#### c. Make a draft of your visualization. (12 points)  

Feel free to be creative with this! _You do not have to do this in R._ You could create a sculpture, painting, textile object, etc.  


If you are making your visualization outside of R, include a photo of your visualization in your submission.  

#### d. Write an artist statement. (8 points)

An artist statement gives the audience context to understand your work. Write 4-5 sentences to address:  

- the content of your piece (what are you showing?)  
- the influences (what did techniques/artists/etc. did you find influential in creating your work?)  
- the form of your work (written code, watercolor, oil painting, etc.)  
- your process (how did you create your work?)  

**For this affective visualization, I aimed to show the diversity of color in the clothes from my closet through a psuedo tile plot. My inspiration for this visualization was to do something fun and personal to me, and to take the idea of a tile plot and make it my own. I chose this project to get a sense of what colors I typically like to wear, and what factors influence these microdecisions. Since coming to University, fashion has been a method of creativity and expression, and I chose this as my project to combine statistics with this passion of mine. I chose this format of visualization to see the diversity in a real life medium, all layed out in front of me. I often forget about clothes in my closet, so doing this visualization reminded me of pieces I love, but don't wear often. My process involved picking my favorite pieces of each color, and laying them out in a color scheme that flowed. I then took a video going across the rows, and took pictures of the tile plot in it's entirety. My space to lay out the clothes was limited, or I would have included even more clothes.**





### Problem 3. Statistical critique (36 points)

At this point, you have seen and created a lot of figures for this class. Revisit the paper you chose for your critique and your homework 2, where you described figures or tables in the text. Address the following in full sentences (3-4 sentences each).  

**For this section of your homework, you will be evaluated on the logic, conciseness, and nuance of your critique.**  

#### a. Revisit and summarize (6 points)

What are the statistical tests the authors are using to address their main research question?  

**The statistical test that this figure describes is a generalized linear model with a binomial distribution. This test is used to test the probability of mortality at three years postfire on first order effects specifically at the Black Mountain Grove. Earlier in the paper, the authors are use a generalized linear mixed model (with a binomial family and logit link) to test the probability of mortality by topographic and tree characteristics.**

Insert the figure or table you described in Homework 2 here.  

#### b. Visual clarity (10 points)

How clearly did the authors _visually_ represent their statistics in figures? For example, are the x- and y-axes in a logical position? Do they show summary statistics (means and SE, for example) and/or model predictions, and if so, do they show the underlying data?  

**The statistics is visually represented well. The models described in table 4 of this paper are confusing, but the corresponding figure is simple and represents the statistics well. The response variable on the y-axis makes sense, and the predictor variable on the x-axis makes sense as well. The underlying data is visible, which makes the comparison of fire scar verses no fire scar clear and understandable. Also, a casual viewer doesn't need to understand the statistics to understand this figure. This figure ties back nicely to the hypothesis, "what first order fire effects best predict mortality by three years postfire at Black Mountain Grove?" by showing that crown damage signficantly predicts mortality of the trees. **

#### c. Aesthetic clarity (10 points)

How well did the authors handle "visual clutter"? How would you describe the the data:ink ratio?  
**The aesthetics of this figure are quite nice. It is simple, clean, and easy to understand. There is no visual clutter, especially when comparing this figure to other figures in the paper. The underlying data is a nice size, adding more information without creating clutter. Every part of the figure is important and necessary, and there is no extraneous information.**


#### d. Recommendations (can be longer than 4 sentences, 10 points)

What recommendations would you make to make the figure better? What would you take out, add, or change? Provide explanations/justifications for each of your recommendations.  

**I would add a title to this figure. Something like, "3 Years Post Fire Data at BLMO." Although there's a description of the figure in the caption, a title would allow for an even easier interpretation. I would also add the actual name of the BLMO acronym (Black Mountain Grove) in the caption. I would do this so a reader doesn't have to go back to the introduction of the paper to understand what grove this figure is actually describing. The only thing I would alter is changing the x-axis from increments of 25 to increments of 20. The paper mentions that 80% crown damage is a significant number, so having it on the x-axis would make it easier to visualize at what specific point does crown damage influence probability of mortality. **

  
### Checklist

Your homework should 

- Include your name, the title (“Homework 3”), and the date you turned in the assignment **(3 points)**  
- Include for Problem 1: 
    - a table for part a (if you made it in code, show the code and annotations) 
    - written response for part b 
    - full work (code and annotations) and output for part c 
    - written response for part d 
    - written response for part e 
- Include for Problem 2: 
    - written response for part a 
    - photo for part b 
    - code and annotations or photo for part c  
    - written response for part d 
- Include for Problem 3: 
    - written responses for part a-d 
    - written responses for parts g-h 
- be uploaded to Canvas as a single PDF **(1 point)** 
- be organized and readable **(10 points)**, meaning that  
    - you are submitting a rendered PDF  
    - all messages and warnings are hidden    
    - all figures and tables are displayed correctly in the rendered PDF   
    - all code annotations are visible in the rendered PDF  
    - you have a set up chunk at the beginning and each section only has the code corresponding to that specific section (for example, no reading in packages/data in the chunk to create a visualization)  
- be associated with a GitHub repository **(15 points)**, meaning that  
    - you have a link to your GitHub repository at the top of your document  
    - you have committed/pushed at least 10 times over the course of completing your homework  
    - at least 10 of your commits/pushes contain a descriptive, concise commit message (a few words describing what changes you are committing)  

**141 points**

